Verifying Android Applications using Java PathFinder

### 2012-09-02 Papers/paper.tex
Paper updates

\numberofauthors{3}

\author{%
\alignauthor Heila van der Merwe\\
       \affaddr{Dept. of Computer Science}\\
       \affaddr{University of Stellenbosch}\\
       \affaddr{Private Bag X1 Matieland}\\
       \affaddr{South Africa, 7602}
       \email{hvdmerwe@cs.sun.ac.za}
\alignauthor  Brink van der Merwe\\
 \affaddr{Dept. of Computer Science}\\
       \affaddr{University of Stellenbosch}\\
       \affaddr{Private Bag X1 Matieland}\\
       \affaddr{South Africa, 7602}
       \email{abvdm@cs.sun.ac.za}
\alignauthor  Willem Visser\\
 \affaddr{Dept. of Computer Science}\\
       \affaddr{University of Stellenbosch}\\
       \affaddr{Private Bag X1 Matieland}\\
       \affaddr{South Africa, 7602}
       \email{wvisser@cs.sun.ac.za}
}

@@

Another approach is to test Android applications using the Java Virtual Machine (JVM). There are different ways of implementing such
a framework. Android lint~\cite{lint}, for example, makes use of static analysis to identify common errors in applications.
Robolectric~\cite{robolectric}, a JUnit testing framework running on the JVM,  intercepts the loading of Android classes and then uses
shadows classes that model these classes.

@@

\section{Design}
\subsection{Java PathFinder}
JPF is an automated, open source, analysis engine for Java applications~\cite{JPFDocs}. It is implemented as an explicit state model checker
that includes mechanisms to model Java classes and native method calls (Model Java Interface - MJI Environment), track byte code execution
and listen for property violations. Additionally, JPF's design encourages developers to create extensions to the framework. Currently there
exist many extensions including a symbolic execution extension (JPF-SYMBC)~\cite{JPF-SYMB}, data race
detector (JPF-RACEFINDER) and an abstract window toolkit (AWT) extension (JPF-AWT)~\cite{JPF-AWT}.

@@

As shown in figure 1, the Android OS is built on top of a modified Linux kernel. The kernel provides a layer of abstraction on top of its
low level functionality such as process, memory, user, network and thread management. On top of the Android kernel is a set of native
C-libraries. The Dalvik VM is a custom, optimised version of the JVM. For security reasons each Android application is run as a separate
process in its own Dalvik VM instance~\cite{AndroidSecurity}.  Hence, applications can only communicate with each other and the with the
Android framework using Android's Binder inter-process communication (IPC) mechanism~\cite{Binder}.

\begin{figure}
\centering
\epsfig{file=Stack.eps, width=3in,}
\caption{The Android application stack~\cite{systemserver}}
\end{figure}

Android has one main application called the system process. The system process contains services responsible for performing the main
tasks of the system including:
\begin{description}
 \item [ActivityManager] manages the life-cycle and interaction of all the activities running on the system
 \item [WindowManager] allows applications to draw on the screen and forwards UI input to the application.
 \item [PackageManager] stores information on the application packages installed on the device
\end{description}

All Android applications follow a single-threaded design in which the main thread of the application handles all application
events~\cite{AndroidDocs}. This structure is commonly used by many UI frameworks since it becomes too complex to make all UI classes thread
safe~\cite{SingleThread}. In Android, this main thread is called the \texttt{Looper} thread. The \texttt{Looper} has a message queue
attached to it containing all application events to be dispatched. UI and system events are scheduled on the main \texttt{Looper} by adding
them to the message queue. The \texttt{Looper} is responsible for continuously looping through these messages and handling them
appropriately. This could include updating a widget, loading a new activity or processing an Intent.

The main entry-point of each Android application is in its \texttt{ActivityThread} class. The \texttt{ActivityThread} class is part of the
Android framework and starts the application's \texttt{Looper} thread. It also keeps track of the application's components and handles user
and system events. Android applications also consists of the following application components:
\begin{description}
\item [Activity] responsible for representing and managing an user interface. An application can consist of many
Activities.
\item [Service] performs background operations such as the pulling of messages from a server every 5 minutes. Services do not have user
interfaces and an application can have zero or more services running simultaneously.
\item [Broadcast receiver] listens for and responds to system-wide events such as network failing, low battery or screen orientation
change events.
\item [Content provider] manages application data stored on the file system, in databases, on the web or other storage medium and provides a
gateway to this data from other applications.
\end{description}

These components interact with each other and with other applications using a structure called an \texttt{Intent}. An \texttt{Intent} is a
high level implementation of the Binder IPC.

@@

JPF-ANDROID focuses on verifying a single application with multiple application components and their interaction. The system service of
the Android OS is not part of the application process and runs in its own thread. To reduce scheduling possibilities, the entire system
service is not modelled but its necessary components are implemented as part of the application process. The following parts of the Android
framework is modelled:
\begin{description}
 \item [ActivityManager]  manages the life-cycle of Activities and other application components.
 \item [ActivityThread]   the main entry-point to the application. It controls and manages the application components and their input.
 \item [Application components] including Activity, Service, Broadcast Receiver and Content Provider. 
 \item [Window and View structure] The view hierarchy is modelled including the widgets and the window classes.
 \item [Message queue] modelled to support input from the script file.
\end{description}

Lastly, as the application will not be communicating with outside processes, the \texttt{Intent} objects are modelled to exclude the
Binder IPC service.

\section{Development}
\subsection{JPF-ANDROID architecture}
\begin{figure}
\centering
\epsfig{file=arch.eps, width=3.5in,}
\caption{JPF-ANDROID architecture}
\end{figure}
As discussed above, Android and AWT applications have a single-threaded application design. That is why JPF-ANDROID's architecture
is based on JPF-AWT. Figure 2 shows JPF-ANDROID's architecture. JPF-ANDROID models Android's message queue structure
by using JPF's MJI environment. When the \texttt{Looper} thread requests a new message from the message queue and it is empty, a call to the
native \texttt{JPF\_MessageQueue} class reads the next event from the input script file.

Android applications can receive input from two different sources: user input and system input. The
\texttt{JPF\_MessageQueue} class classifies events as either UI or system events. UI events are directed to and handled by the native
\texttt{JPF\_Window} class and system events are handled by the \texttt{ActivityManager} implemented as the
native \texttt{JPF\_ActivityManager} class.

When an Activity's UI (window) is inflated, an object map is created in the native \texttt{JPF\_Window} class. This
map binds the name of each widget to the reference of the inflated widget object. When an UI event is received by the native
\texttt{JPF\_Window} class, the name of the target widget is looked-up in the object map and the action is then called directly on the
inflated widget object by pushing a direct call frame on the JPF call stack.

System events are a little more complex. They use \texttt{Intent} objects to describe a system event. \texttt{Intents} are used to start
Activities and Services or to provide a notification of certain events. They are similar to messages containing a description of an
operation to be performed or, often in the case of broadcasts, a description of something that has happened and is being
announced~\cite{AndroidDocs}. System events are handled by the native \texttt{JPF\_ActivityManager} class. This class keeps a map of Intents
as they are created in the script file. When a system event is received, the \texttt{ActivityManager} looks up the corresponding
\texttt{Intent} in the map. It then resolves the \texttt{Intent} to the relevant application and sends it to the application's
\texttt{ActivityThread} to be handled. This message will be processed by the \texttt{Looper} which will direct the \texttt{Intent} to the
corresponding application component.

\subsection{The input script}
AWT applications that are tested with JPF-AWT contains at least one input script. This script consist of a list of UI events. These
events are executed one-by-one when the message queue it empty. JPF-AWT's syntax only supports the scripting of UI events. To
allow users to script system events, variables were added to the scripting language. Variable names are identified by the ``@'' in front of
the name as '\$' is already used in JPF-AWT to identify UI components. These variable can be used to construct most Intent objects. For
example, a script file that sends an \texttt{Intent} to start the \texttt{SampleActivity} will contain:

{\small
{\sf
\hspace*{4mm}@intent1.setComponent(``com.example.SampleActivity'')\\
\hspace*{4mm}startActivity(@intent1)\\
\hspace*{4mm}\$button1.onClick()
}
}

JPF-ANDROID scripts also adopted the REPEAT and ANY constructs from JPF-AWT. REPEAT constructs repeat a list of events a specified number
of times. ANY constructs take a list of events and then uses a ChoiceGenerator and JPF's state matching and backtracking features to visit
each of the execution branches. JPF stores the state of the system before the it advances to a new state. JPF-AWT uses a JPF search listener
to store and retrieve the current position of the input script in a specific state so that the script's state is also saved.

Android applications contain multiple windows - one for each Activity. This complicates the script input as each window has its own unique
set of view components, hence, a unique input sequence. Furthermore, the application flow can switch between these Activities
at any point of execution. In other words, if we do not know in what Activity the application is, we can not schedule
the following events.

\begin{figure}
\centering
\epsfig{file=newContact.eps, width=1in,}\\
\caption{Contacts application window}
\end{figure}
\begin{figure}
{\small
{\sf
...\\
\textbf{ANY} \{ \$cancelButton.onclick(), \$clearButton.onClick()  \}
\$nameEdit.setText(``Mary'')\\
\$createButton.onClick()\\
...
}
}
\caption{Input script for ``New Contact'' window}
\end{figure}
Figure 3 displays the ``New Contact'' window of a Contacts application. It contains three
buttons: the cancel button restarts the \texttt{ListContactsActivity}, the clear button stays on the current Activity and the create 
button starts the \texttt{ViewContactActivity}. Now let us look at figure 4 containing the input sequence for the ``New Contact'' window.

The problem occurs in the ANY structure. When the cancelButton.onclick() event fires the application changes to another
Activity and the following events are not valid any more. To address this issue, we included the
use of sections in the input script (see figure 5). Each section groups the input events of a specific Activity. Now, if the cancel button
is pressed the \texttt{ListContactsActivity} will be started and then \texttt{ListContactsActivity}'s event sequence
will execute instead of executing the following events.

\begin{figure}
{\small
{\sf 
\textbf{SECTION} default \{\\
\hspace*{5mm}@startIntent.setComponent(``com.example.ListContactsActivity'')\\
\hspace*{5mm}startActivity(@startIntent)\\
\}

\textbf{SECTION} com.example.ListContactsActivity \{\\
\hspace*{5mm}\$createButton.onClick()\\
\hspace*{5mm}\$list.setSelectedIndex([0-3])\\
\}\\

\textbf{SECTION} com.example.AddContactActivity \{\\
\hspace*{5mm}\textbf{ANY} \{ \$cancelButton.onclick(), \$clearButton.onClick()  \}
\hspace*{5mm}\$nameEdit.setText(``Mary'')\\
\hspace*{5mm}\$createButton.onClick()\\
\}
}
}
\caption{Adding sections to the input script}
\end{figure}

Now lets view an Android application as a graph with each Activity as its nodes. The root of the tree is the applications main Activity. The
next challenge is to avoid an infinite loops in the application flow. If we look at the input in figure 5, an infinite loop will occur. The
ListContactsActivity starts the AddContactActivity and when the cancel button is pressed the AddContactActivity returns to
the ListContactsActivity. To solve this problem, JPF-ANDROID stores its current Activity. When input events are requensted by the input
script they are read from the current Activity's script section. The current position in this script file is also stored. When the
application forwards to a next activity, the state of all the visited activities in the execution path is stored. So when JPF backtracks
the the state of previous Activity's script files are continued instead of restarted. So when AddContactActivity returns to
ListContactsActivity, the script is not restarted but continued.

\subsection{Case study}

Currently JPF-ANDROID can detect deadlocks, race conditions and other property violations in Android applications. The challenge is
providing a sample application to show how JPF-ANDROID works. This is because most Android applications make use of Android specific
libraries such as a sqlite database connector, HTTP connection or the mediaplayer libraries. The future work of JPF-ANDROID
includes modelling these Android specific libraries.

To show how JPF-ANDROID works, we will use a calculator application.

@@

This extension provides a basis on which Android applications can be tested. It can later be extended to verify functional requirements
and identify Android specific errors using JPF's listener mechanism.

Future extensions include modelling a sqlite database and http connections.

%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{paper}  % sigproc.bib is the name of the Bibliography in this case

>>>>

\numberofauthors{1}

\author{%
\alignauthor Heila van der Merwe, Brink van der Merwe and Willem Visser\\
       \affaddr{Dept. of Computer Science}\\
       \affaddr{University of Stellenbosch}\\
       \affaddr{Private Bag X1 Matieland}\\
       \affaddr{South Africa, 7602}
       \email{hvdmerwe@cs.sun.ac.za, abvdm@cs.sun.ac.za, wvisser@cs.sun.ac.za }
}

@@

Another approach is to test Android applications using the Java Virtual Machine (JVM). There are different ways of implementing such
a framework. Android lint, for example, makes use of static analysis to identify common errors in applications.
Robolectric~\cite{robolectric}, a JUnit testing framework running on the JVM,  intercepts the loading of Android classes and then uses
shadows classes that model these classes.

@@

\section{Design}
\subsection{Java PathFinder}
JPF is an automated, open source, analysis engine for Java applications~\cite{JPFDocs}. It is implemented as an explicit state model checker
that includes mechanisms to model Java classes and native method calls (Model Java Interface - MJI Environment), track byte code execution
and listen for property violations. Additionally, JPF's design encourages developers to create extensions to the framework. Currently there
exist many extensions including a symbolic execution extension (JPF-SYMBC), data race
detector (JPF-RACEFINDER) and an abstract window toolkit (AWT) extension (JPF-AWT)~\cite{JPF-AWT}.

@@

As shown in figure \ref{fig:android}, the Android OS is built on top of a modified Linux kernel. The kernel provides a layer of abstraction
on top of its low level functionality such as process, memory, user, network and thread management. On top of the Android kernel is a set of
native C-libraries. The Dalvik VM is a custom, optimised version of the JVM. For security reasons each Android application is run as a
separate process in its own Dalvik VM instance~\cite{AndroidSecurity}.  Hence, applications can only communicate with each other and the
with the Android framework using Android's Binder inter-process communication (IPC) mechanism~\cite{Binder}.

\begin{figure}
\centering
\epsfig{file=Stack.eps, width=3in,}
\caption{The Android application stack~\cite{systemserver}}
\label{fig:android}
\end{figure}

Android has one main application called the system process. The system process contains services responsible for performing the main
tasks of the system including:
\vspace{-10pt}
\begin{description}
 \item [ActivityManager] manages the life-cycle and interaction of all the activities running on the system
 \item [WindowManager] allows applications to draw on the screen and forwards UI input to the application.
 \item [PackageManager] stores information on the application packages installed on the device
\end{description}
\vspace{-5pt}
All Android applications follow a single-threaded design in which the main thread of the application handles all application
events~\cite{AndroidDocs}. This structure is commonly used by many UI frameworks since it becomes too complex to make all UI classes thread
safe~\cite{SingleThread}. In Android, this main thread is called the \texttt{Looper} thread. The \texttt{Looper} has a message queue
attached to it containing all application events to be dispatched. UI and system events are scheduled on the main \texttt{Looper} by adding
them to the message queue. The \texttt{Looper} is responsible for continuously looping through these messages and handling them
appropriately. This could include updating a widget, loading a new activity or processing an Intent.

The main entry-point of each Android application is in its \texttt{ActivityThread} class. The \texttt{ActivityThread} class is part of the
Android framework and starts the application's \texttt{Looper} thread. It also keeps track of the application's components and handles user
and system events. Android applications also consists of the following application components:
\vspace{-10pt}
\begin{description}
\item [Activity] responsible for representing and managing an user interface. An application can consist of many
Activities.
\item [Service] performs background operations such as the pulling of messages from a server every 5 minutes. Services do not have user
interfaces and an application can have zero or more services running simultaneously.
\item [Broadcast receiver] listens for and responds to system-wide events such as network failing, low battery or screen orientation
change events.
\item [Content provider] manages application data stored on the file system, in databases, on the web or other storage medium and provides a
gateway to this data from other applications.
\end{description}
\vspace{-5pt}
These components interact with each other and with other applications using a structure called an \texttt{Intent}. An \texttt{Intent} is a
high level implementation of the Binder IPC. \texttt{Intents} are used to start Activities and Services or to provide a notification of
certain events. They are similar to messages containing a description of an operation to be performed or, often in the case of broadcasts, a
description of something that has happened and is being announced~\cite{AndroidDocs}.

@@

JPF-ANDROID focuses on verifying a single application with multiple application components and their interaction. The system service of
the Android OS is not part of the application process and runs in its own thread. To reduce scheduling possibilities, the entire system
service is not modelled but its necessary components are implemented as part of the application process. The following parts of the Android
framework is modelled:
\vspace{-10pt}
\begin{description}
 \item [ActivityManager]  manages the life-cycle of Activities and other application components.
 \item [ActivityThread]   the main entry-point to the application. It controls and manages the application components and their input.
 \item [Application components] including Activity, Service, Broadcast Receiver and Content Provider. 
 \item [Window and View structure] The view hierarchy is modelled including the widgets and the window classes.
 \item [Message queue] modelled to support script input.
\end{description}
\vspace{-5pt}
Lastly, as the application will not be communicating with outside processes, the \texttt{Intent} objects are modelled to exclude the
Binder IPC service.

\section{Development}
\subsection{JPF-ANDROID architecture}
% \begin{figure}
% \centering
% \epsfig{file=arch.eps, width=3.5in,}
% \caption{JPF-ANDROID architecture}
% \label{fig:arch}
% \end{figure}
%JPF-ANDROID BASED ON JPF_AWT  why and how does it work?
As discussed above, both Android and AWT applications have a single-threaded application design. That is why JPF-ANDROID's architecture
is based on JPF-AWT. JPF-ANDROID models Android's message queue by using JPF's MJI
environment. When the \texttt{Looper} thread requests a new message from the message queue and it is empty, a call to the native
\texttt{JPF\_MessageQueue} class requests the next event from the scripting environment. The \texttt{JPF\_MessageQueue} class classifies
events as either UI or system events. UI events are directed to and handled by the native \texttt{JPF\_Window} class and system events by
the native \texttt{JPF\_ActivityManager} class.

% Explain what UI events are and how they are handled by android vs how they ae handled by the system
UI events include events fired by widgets such as clicking a button or selecting an item in a list view. An
Activity's UI is represented by a window object containing a view hierarchy. When the window is inflated, an object map is created in the
native \texttt{JPF\_Window} class. This map binds the name of each widget to the reference of the inflated widget object. When an UI event
is received by the native \texttt{JPF\_Window} class, the name of the target widget is looked-up in the object map and the action is then
called directly on the inflated widget object by pushing a direct call frame on the JPF call stack.

% Explain what system events are how they are handled by android how they are handled by the system
System events use \texttt{Intent} objects to describe a the event that occurred. They include events to start an Activity or Service or a
battery low notification. In the Android OS, system events are sent to the Activity manager which is modelled in JPF-ANDROID by the native
\texttt{JPF\_ActivityManager} class. This class keeps a map of Intents as they are defined in the script file. When a system event is
received, the corresponding \texttt{Intent} is looked up in the map. The \texttt{Intent} is then resolved to the relevant
component and scheduled in the application's message queue.

\subsection{The scripting environment}
% what is input script, how does it work(its function) and how are does it look
JPF-AWT's script consist of a list of UI events. These events are executed one-by-one when the message queue it empty. To allow users to
script system events, variables were added to the scripting language in JPF-ANDROID. Variable names are identified by the ``@'' in front of
the variable's name as '\$' is already used in JPF-AWT to identify UI components. These variable can be used to construct most Intent
objects. For example, a script file that sends an \texttt{Intent} to start the \texttt{SampleActivity} is shown in figure \ref{fig:start}
\begin{figure}
\frame{
\parbox{0.48\textwidth}{
{\small
{\sf 
\vspace*{2mm}
\hspace*{2mm}1. @intent1.setComponent(``SampleActivity'')\\
\hspace*{2mm}2. startActivity(@intent1)\\
\hspace*{2mm}3. \$button1.onClick()
\vspace*{2mm}
}}}}
\vspace{-18pt}
\caption{Starting SampleActivity}
\label{fig:start}
\end{figure}

JPF-ANDROID scripts also adopted the \textbf{REPEAT} and \textbf{ANY} constructs from JPF-AWT. \textbf{REPEAT} constructs repeat a list of
events a specified number of times. \textbf{ANY} constructs take a list of events as parameters and then uses a ChoiceGenerator and JPF's
state matching and backtracking features to visit each of the execution branches. JPF stores the state of the system before the it advances
to a new state. JPF-AWT uses a JPF search listener to store and retrieve the current position of the input script in a specific state so
that the script's state is also saved.

%how and why was this scripting environment adapted
Android applications contain multiple windows - one for each Activity. This complicates the script input as each window has its own unique
set of view components, hence, a unique input sequence. Furthermore, the application flow can switch between these Activities
at any point of execution. In other words, if we do not know in what Activity the application is, we can not script the following events.

% \begin{figure}
% \centering
% \epsfig{file=newContact.eps, width=1in,}\\
% \caption{Contacts application window}
% \end{figure}
% \begin{figure}
% {\small
% {\sf
% ...\\
% \textbf{ANY} \{ \$cancelButton.onclick(), \$clearButton.onClick()  \}
% \$nameEdit.setText(``Mary'')\\
% \$createButton.onClick()\\
% ...
% }
% }
% \caption{Input script for ``New Contact'' window}
% \end{figure}
Let us take a contacts android application as an example. The application has two Activities: ListContactsActivity and the
NewContactActivity. The ListContactsActivity contains a list of the current contacts and a button to add new contacts that directs the user
to the NewContactActivity. The NewContactActivity contains two buttons: the clear button stays on the current Activity and clears its fields
and the create button starts the \texttt{ListContactsActivity}. Now let us look at figure \ref{fig:contact} containing the input sequence
for the NewContactActivity.

\begin{figure}
\frame{
\parbox{0.48\textwidth}{
{\small
{\sf 
\vspace*{2mm}
\hspace*{2mm}1. @startIntent.setComponent(``ListContactsActivity'')\\
\hspace*{2mm}2. startActivity(@startIntent)\\
\hspace*{2mm}3. \$createContactButton.onClick()\\
\hspace*{2mm}4. \$nameEdit.setText(``Mary'')\\
\hspace*{2mm}5. \textbf{ANY} \{ \$<create\,|\,clear>Button.onClick()\}\\
\hspace*{2mm}6. \$nameEdit.setText(``Maria'')\\
}}}}
\vspace{-15pt}
\caption{Original script for Contacts application}
\label{fig:contact}
\end{figure}

The problem occurs in the ANY structure. When the create button is clicked the application changes to the ListContactsActivity
and the following event is not valid any more. This drastically restricts the number of sequences that can be scripted in the file. To
address this issue, we included the use of sections in the input script (see figure \ref{fig:contact2}). Each section groups the input
events of a specific Activity. Now, if the create button is pressed the \texttt{ListContactsActivity} will be started and the
\texttt{ListContactsActivity}'s events specified in its sequence will start executing again.

\begin{figure}
\frame{
\parbox{0.48\textwidth}{
{\small
{\sf 
\vspace*{2mm}
\hspace*{2mm}1. \textbf{SECTION} default \{\\
\hspace*{2mm}2. \hspace*{5mm}@startIntent.setComponent(``ListContactsActivity'')\\
\hspace*{2mm}3. \hspace*{5mm}startActivity(@startIntent)\\
\hspace*{2mm}4. \}\\
\hspace*{2mm}5. \\
\hspace*{2mm}6. \textbf{SECTION} ListContactsActivity \{\\
\hspace*{2mm}7. \hspace*{5mm}\$createButton.onClick()\\
\hspace*{2mm}8. \hspace*{5mm}\$list.setSelectedIndex([0-3])\\
\hspace*{2mm}9. \}\\
\hspace*{2mm}10. \\
\hspace*{2mm}11. \textbf{SECTION} NewContactActivity \{\\
\hspace*{2mm}12. \hspace*{5mm}\$nameEdit.setText(``Mary'')\\
\hspace*{2mm}13. \hspace*{5mm}\textbf{ANY} \{ \$<create\,|\,clear>Button.onClick()\}\\
\hspace*{2mm}14. \hspace*{5mm}\$nameEdit.setText(``Maria'')\\
\hspace*{2mm}15. \}
\vspace*{2mm}
}}}}
\vspace{-15pt}
\caption{Adding sections to the input script}
\label{fig:contact2}
\end{figure}

The addition of the sections in the input script lead to infinite event sequences. In the above example, an infinite loop occurs when the
create button is pressed on the NewContactActivity window. This will stop the execution of the NewContactActivity section and restart
the ListContactsActivity section's events. When the ListContactsActivity's section is restarted, it again clicks on the createButton
restarting the NewContactActivity. To address this issue, JPF-ANDROID keeps record of its current Activity. The scripting environment was
also adapted to store the current position of each section's events for each visited Activity. If the branch returns to a previously visited
Activity, its current position in its section is looked up and instead of restarting the section, it continues from its previous position.

\subsection{Case studies}
%calculator
%how does application look and work
The first application is a scientific calculator. The calculator has two Activities: a simple view that
displays basic arithmetic operations and a scientific view that displays more complex arithmetic operations. When the user switches
between these Activities, the current state of the calculator is preserved. This state includes the intermediate values and current
operation. This state information is bundled with the \texttt{Intent} that starts then next Activity.
\begin{figure}

\frame{
\parbox{0.48\textwidth}{
{\small
{\sf 
\vspace*{2mm}
\hspace*{2mm}1. \textbf{SECTION} SimpleActivity  \{\\
\hspace*{2mm}2. \hspace*{5mm}\textbf{ANY} \{\,\$button[0-9].onClick()\,\}\\
\hspace*{2mm}3. \hspace*{5mm}\textbf{ANY} \{\,\$button<Plus|Minus|Mul|Div|More>.onClick()\,\}\\
\hspace*{2mm}4. \hspace*{5mm}\textbf{ANY} \{\,\$button[0-9].onClick()\,\}\\
\hspace*{2mm}5. \hspace*{5mm}\$buttonEquals.onClick()\\
\hspace*{2mm}6. \}\\
\hspace*{2mm}7. \textbf{SECTION} ScientificActivity  \{\\
\hspace*{2mm}8. \hspace*{5mm}\textbf{ANY} \{\,\$button<Sin|Cos|Tan>.onClick()\,\}\\
\hspace*{2mm}9. \}\\
}}}}
\vspace{-15pt}
\caption{Input script for Calculator application}
\label{fig:calc}
\end{figure}

%what errors can occur
The calculator application contains two errors. When the user divides a value by zero, an uncaught
\texttt{ArithmeticException} is thrown by the application which causes Android to kill the application. Secondly the application neglected
to attach the state information to the Intent that is passed to the next Activity. When the user switches to the other Activity a
\texttt{NullPointerException} is thrown when it tries to read the state information from the Intent.

To detect these errors we will use the test script in figure~\ref{fig:calc}. The scripting environment interprets the script into the
following input sequences:
\begin{align*}
\text{<sequence>} &= \text{<simple\_seq>}\,|\,\text{<complex\_seq>}\\
\text{<simple\_seq>} &= \text{number},\,\text{simple\_op},\,\text{number},\,\text{equals}\\
\text{<complex\_seq>} &= \text{number},\,\text{complex\_op}\\
\text{simple\_op} &= \text{``+''}\,|\,\text{``-''}\,|\,\text{``}\times\text{''}\,|\,\text{``}\div\text{''}\\
\text{complex\_op} &= \text{``sin''}\,|\,\text{``cos''}\,|\,\text{``tan''}\\
\text{equals} &= \text{``=''} \\
\text{number} &= \text{``0''\,|\,``1''\,|\,``2''\,|\,``3''\,|\,``4''\,|\,``5''\,|\,``6''\,|\,``7''\,|\,``8''\,|\,``9''}
\end{align*}
Each of the ten number buttons are pressed followed by a simple operation and another number or a complex operation.

%how they where detected
JPF is automatically configured to detect thrown exceptions and to stop execution. When JPF-ANDROID was run it firstly detected the
\texttt{ArithmenticException} due to the division by zero:
\begingroup
 \fontsize{6pt}{7pt}\selectfont
\begin{verbatim}
====================================================== results
error #1: gov.nasa.jpf.jvm.NoUncaughtExceptionsProperty 
         "java.lang.ArithmeticException: Division by zero  at ..."
====================================================== statistics
elapsed time:       00:00:01
states:             new=76, visited=0, backtracked=63, end=30
search:             maxDepth=13, constraints hit=0
choice generators:  thread=7 (signal=0, lock=3, shared ref=0), data=39
heap:               new=3455, released=2105, max live=1507, gc-cycles=74
instructions:       76796
max memory:         117MB
loaded code:        classes=145, methods=2064
======================================================
\end{verbatim}
\endgroup
After this error was fixed JPF-ANDROID detected the \texttt{NullPointerException} in the same way.

Both of these errors would have been difficult to detect with unit testing. The \texttt{ArithmenticException} is challenging due
to the many possible input sequence combinations. If a test case did not specifically identify this as a point of interest, unit testing
would not have detected this error. The second error is challenging to detect due to the fact that it only occurs when the flow of Activity
classes are modelled.

% deadlock
%how does application look and work
The next case study is a very simple application demonstrating how JPF-ANDROID detects a deadlock in a Android application. When the
\texttt{Looper}/main thread of an application is caught in a deadlock, the Android OS kills the application and displays an
Application Not Responding (ANR) dialog. But, Android does not detect a deadlock if it occurs between other asynchronous threads. This
sample application spawns two asynchronous threads that deadlock. The application has one Activity with two
buttons. The first button spawns the first thread and the second button spawns the second thread. After a while these two thread deadlock
and are then blocked forever, waiting for each other. The input script for the application is given in figure \ref{fig:deadlock}.
\begin{figure}
\frame{
\parbox{0.48\textwidth}{
{\small
{\sf 
\vspace*{2mm}
\hspace*{2mm}1. \textbf{SECTION} DeadlockActivity  \{\\
\hspace*{2mm}2. \hspace*{5mm}\$button1.onClick()\\
\hspace*{2mm}3. \hspace*{5mm}\$button2.onClick()\\
\hspace*{2mm}4. \}\\
}}}}
\vspace{-15pt}
\caption{Input script for Deadlock application}
\label{fig:deadlock}
\end{figure}

JPF is then configured to listen for deadlocks and schedules the threads in all possible ways to detect the deadlock.
\begingroup
 \fontsize{6pt}{7pt}\selectfont
\begin{verbatim}
 ====================================================== thread ops #1
   1       1     trans          loc            : stmt
------- ------- ---------------------------------------------------
B:1003     |      54  DeadlockActivity.java:82 : bower.bowBack(this);
   |    B:1000    54  DeadlockActivity.java:82 : bower.bowBack(this);
L:1000     |      54  DeadlockActivity.java:56 : friend[0].bow(friend[1]);
   |    L:1003    18  DeadlockActivity.java:56 : friend[0].bow(friend[1]);
   S       |       6
   |       S       3
====================================================== results
error #1: gov.nasa.jpf.jvm.NotDeadlockedProperty 
         "deadlock encountered:   thread java.lang.Thread:{i..."
====================================================== statistics
elapsed time:       00:00:01
states:             new=55, visited=13, backtracked=67, end=1
search:             maxDepth=10, constraints hit=0
choice generators:  thread=26 (signal=0, lock=11, shared ref=0), data=7
heap:               new=1430, released=375, max live=1040, gc-cycles=67
instructions:       12661
max memory:         117MB
loaded code:        classes=140, methods=1792
======================================================
\end{verbatim}
\endgroup

\section{Future Work}
Currently JPF-ANDROID can detect deadlocks, race conditions and other property violations in Android applications. The next challenge is
modelling the extra Android libraries. This is because most Android applications make use of many Android specific
libraries such as the sqlite database connector, HTTP connections or the media player libraries.

Another extension that will be added to JPF-ANDROID is coverage testing. JPF has many coverage extensions available so we will be adapting
one these extensions to work on Android applications.

@@

This extension provides a basis on which Android applications can be tested. It can later be extended to verify functional requirements
and identify Android specific errors using JPF's listener mechanism.
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\scriptsize
\bibliography{paper}  % sigproc.bib is the name of the Bibliography in this case
